---
title: "DATA612 Final Project"
author: "Michael O'Donnell"
date: "July 10, 2019"
output: html_document
---

# Overview: 

## Dataset: 157,500 book reviews by 700 users
## User Recommendations: Binary, 1 for 'liked' and 0 for 'did not like'
## Recommender System: User-Based Collaborative Filtering system

### Loading the necessary libraries
```{r, results = 'hide'}
library(data.table)
library(ggplot2)
library(recommenderlab)
library(tibble)
library(dplyr)
library(readr)
set.seed(1)
```

### Reading in the dataset as a CSV file
```{r, results = 'hide'}
book_reviews <- read_csv("book_reviews.csv", col_types =  cols(
  .default = col_integer(),
  user = col_character()
))
```

### most frequently liked book
```{r}
most_liked <- which.max(colMeans(book_reviews[2:288]))
most_liked
```

### least frequently liked book
```{r}
least_liked <- which.min(colMeans(book_reviews[2:288]))
least_liked
```

### Replacing the 0's and 1's with TRUE and FALSE, in preparation for Binary Rating Matrix
```{r, results = 'hide'}
ratings <- book_reviews %>% 
  # now we replace all NAs by 0
  mutate_each(funs(replace(., is.na(.), 0))) %>%
  # now we convert all 0 to false and 1 to true
  # (except the user column, because it contains names and neighter 0 nor 1)
  mutate_each(funs(as.logical), -user) %>% 
  # dplyr uses a own structure called tibble, but we want dataframes
  # .. so we convert it to one
  as.data.frame() %>%
  # .. and we convert a column to the rowname of the dataframe
  column_to_rownames(var ="user") 
```

### Creating a Binary Rating Matrix
```{r, results = 'hide'}
binaryBookRatings <- as.matrix(ratings) %>% as("binaryRatingMatrix")
```

### View the size of the book review data
```{r}
object.size(binaryBookRatings)
object.size(as(binaryBookRatings, "matrix"))
```

### Setting up the UBCF recommender model
```{r, results = 'hide'}
model <- Recommender(data = binaryBookRatings, method = "UBCF", 
                     parameter = list(method = "cosine"))
```

### Creating 1 book recommednations for each user
```{r echo = T, results = 'hide'}
recommendations <- predict(model, binaryBookRatings, n = 1)
```

### See recommedations for all users
```{r}
#getList(recommendations)
```

### see recommended book for 1 random test user
```{r}
user_one <- recommendations@items[[sample(1:700, 1)]]
test_book_one <- recommendations@itemLabels[user_one]
cat("User", user_one, ":", test_book_one)
```

### Turn all recommendations into dataframe
```{r echo = T, results = 'hide'}
book_recs <- as(recommendations, "matrix") %>% as.data.frame()
book_recs[is.na(book_recs)] <- 0
```

### Find the top 5 recommended books by the UBCF system
```{r}
colCount = colSums(book_recs[-1])
topFiveIds = order(colCount,decreasing=TRUE)[1:5] + 1 # From Joris
topFiveCols = names(book_recs[topFiveIds])
topFiveCols
```
